### 2018-05-29 submission/manuscript.Rmd
redo third section of results

***The number of OTUs are dependent on polymerase:***
Differences in the range of the number of OTUs detected for fecal samples is dependent on the polymerase used (e.g. Accuprime at 35 cycles = `r filter(fecal_numOTU_data, cycles == "35x", taq == "Accuprime") %>% pull(taq_min)` - `r filter(fecal_numOTU_data, cycles == "35x", taq == "Accuprime") %>% pull(taq_max)` versus Phusion at 35 cycles = `r filter(fecal_numOTU_data, cycles == "35x", taq == "Phusion") %>% pull(taq_min)` - `r filter(fecal_numOTU_data, cycles == "35x", taq == "Phusion") %>% pull(taq_max)`) [Figure 1]. Additionally, there is a trend for lower number of cycles (15-20) to result in a reduced range in the number of OTUs versus higher number of cycles (25, 30, and 35) for all polymerases (e.g. Phusion at 15 cycles = `r filter(fecal_numOTU_data, cycles == "15x", taq == "Phusion") %>% pull(taq_min)` - `r filter(fecal_numOTU_data, cycles == "15x", taq == "Phusion") %>% pull(taq_max)` versus Phusion at 35 cycles = `r filter(fecal_numOTU_data, cycles == "35x", taq == "Phusion") %>% pull(taq_min)` - `r filter(fecal_numOTU_data, cycles == "35x", taq == "Phusion") %>% pull(taq_max)`) [Figure 1]. There is an overall difference in the number of OTUs detected within fecal samples between polymerases at 35 cycles (F-stat > `r round(min(filter(fs_anova_summary, cycle == "35x", sub_sample_level == 1000)[, "F.value"]), digits = 2)`, P-value = `r format(filter(fs_anova_summary, cycle == "35x", sub_sample_level == 1000)[, "bh"], digits = 2, nsmall = 2, scientific = T)`) [Table S1]. Using a Tukey post-hoc test to identify which polymerase groups were different from each other at 35 cycles, no difference was found (P-value `r ifelse(min(filter(fs_tukey_summary, sub_sample_level == 1000)[, "padj"]) > 0.05,  invisible(">"), invisible("check data"))` 0.05) [Table S2]. The polymerase dependent difference in the range of the number of OTUs also was observed in the mock community samples [Figure 2]. The closest the polymerases came to the total of `r pull(ref_mock_otu_num, numOtus)` OTUs created by the mock reference 16S sequences was at 25 and 30 cycles [Figure 2]. Regardless of if fecal or mock communities were used, the same polymerases generated high and low number of OTUs and this was consistent across the number of cycles used [Figure 1-2 & Table S1-S2].  In contrast to the results obtained with fecal samples, differences between polymerases for the number of OTUs created were observed as early as 20 cycles in the mock community (F-stat = `r round(filter(mock_anova_summary, sub_sample_level == 1000, cycle == "20x")[, "F.value"], digits = 2)`, P-value = `r format(filter(mock_anova_summary, sub_sample_level == 1000, cycle == "20x")[, "bh"], digits = 1, nsmall = 1)`) [Table S1]. Using a Tukey post-hoc test, the majority of differences for the number of OTUs detected in the mock community is largely due to Kappa and Platinum differences versus the other polymerases across the different number of cycles [Table S2]. Based on these observations in fecal and mock communities, it is clear that using different polymerases result in a different total number of OTUs within a sample.

@@

***Sequence Error is Dependent on both Polymerase and Cycle Number:***
Differences in the median error rate varied by polymerase, with Kappa being the highest [Figure 4 & Table S3]. The differences in the median error rate between the polymerases also is dependent on the number of cycles, with Platinum and Kappa having the most differences versus other polymerases [Figure 4 and Table S4]. The total sequences with at least one error also is dependent on the number of cycles [Table S3 & S4]. These differences in sequences with at least one error is mostly due to differences in Accuprime^TM^ and Kappa versus the other polymerases [Table S3 & S4]. The differences in error rates were not due to polymerase dependent differences in base substitution rate [Figure S1]. These results suggest that sequence error is dependent on polymerase and the number of cycles, but is not attributable to a specific type. Based on these results, we examined whether chimeras also were dependent on polymerase and whether this also could affect the number of OTUs.

***Prevalence of Chimeric Sequences are Polymerase Dependent and Correlate with the Number of OTUs:***
There is significant differences in the chimera prevalence based on polymerase at all the number of cycles used (P-value `r ifelse(max(filter(kruskal_chimera_prev)[, "bh"]) < 0.05, invisible("<"), invisible("check data"))` 0.05) [Table S3]. Differences in chimera prevalence between Platinum and all other polymerases accounted for the majority of these differences [Table S4]. Accuprime^TM^ had the lowest chimera prevalence of all polymerases regardless of whether chimera removal with VSEARCH was used [Figure 5A & 5B]. Additionally, there was a plateau in the total percent of chimeras that were removed that was similar for all polymerases [Figure 5C]. A positive correlation was observed between chimeric sequences and the number of OTUs for all polymerases [Figure 6]. This positive correlation was strongest for Accuprime^TM^, Platinum, and Phusion [Figure 6]. This data suggests that chimera prevalence depends on polymerase used and confirms that the number of OTUs is dependent on the prevalence of these chimeric sequences.

>>>>

***The number of OTUs generated are dependent on polymerase used:***
Differences in the range of the number of OTUs detected for fecal samples is dependent on the polymerase used (e.g. Accuprime at 35 cycles = `r filter(fecal_numOTU_data, cycles == "35x", taq == "Accuprime") %>% pull(taq_min)` - `r filter(fecal_numOTU_data, cycles == "35x", taq == "Accuprime") %>% pull(taq_max)` versus Phusion at 35 cycles = `r filter(fecal_numOTU_data, cycles == "35x", taq == "Phusion") %>% pull(taq_min)` - `r filter(fecal_numOTU_data, cycles == "35x", taq == "Phusion") %>% pull(taq_max)`) [Figure 1]. Additionally, there is a trend for lower number of cycles (15-20) to result in a reduced range in the number of OTUs versus higher number of cycles (25, 30, and 35) for all polymerases (e.g. Phusion at 15 cycles = `r filter(fecal_numOTU_data, cycles == "15x", taq == "Phusion") %>% pull(taq_min)` - `r filter(fecal_numOTU_data, cycles == "15x", taq == "Phusion") %>% pull(taq_max)` versus Phusion at 35 cycles = `r filter(fecal_numOTU_data, cycles == "35x", taq == "Phusion") %>% pull(taq_min)` - `r filter(fecal_numOTU_data, cycles == "35x", taq == "Phusion") %>% pull(taq_max)`) [Figure 1]. There is an overall difference in the number of OTUs detected within fecal samples between polymerases at 35 cycles (F-stat > `r round(min(filter(fs_anova_summary, cycle == "35x", sub_sample_level == 1000)[, "F.value"]), digits = 2)`, P-value = `r format(filter(fs_anova_summary, cycle == "35x", sub_sample_level == 1000)[, "bh"], digits = 2, nsmall = 2, scientific = T)`) [Table S1]. Using a Tukey post-hoc test to identify which polymerase groups were different from each other at 35 cycles, no difference was found (P-value `r ifelse(min(filter(fs_tukey_summary, sub_sample_level == 1000)[, "padj"]) > 0.05,  invisible(">"), invisible("check data"))` 0.05) [Table S2]. The polymerase dependent difference in the range of the number of OTUs also was observed in the mock community samples [Figure 2]. The closest the polymerases came to the total of `r pull(ref_mock_otu_num, numOtus)` OTUs created by the mock reference 16S sequences was at 25 and 30 cycles [Figure 2]. Regardless of if fecal or mock communities were used, the same polymerases generated high and low number of OTUs and this was consistent across the number of cycles used [Figure 1-2 & Table S1-S2].  In contrast to the results obtained with fecal samples, differences between polymerases for the number of OTUs created were observed as early as 20 cycles in the mock community (F-stat = `r round(filter(mock_anova_summary, sub_sample_level == 1000, cycle == "20x")[, "F.value"], digits = 2)`, P-value = `r format(filter(mock_anova_summary, sub_sample_level == 1000, cycle == "20x")[, "bh"], digits = 1, nsmall = 1)`) [Table S1]. Using a Tukey post-hoc test, the majority of differences for the number of OTUs detected in the mock community is largely due to Kappa and Platinum differences versus the other polymerases across the different number of cycles [Table S2]. Based on these observations in fecal and mock communities, it is clear that using different polymerases result in a different total number of OTUs within a sample.

@@

***Sequence error varies by polymerase and is consistent across the number of cycles used:***
The median error rate varied by polymerase, with Kappa having the highest error rate of all the polymerases across the number of cycles [Figure 4 & Table S3]. The majority of the differences across the number of cycles was between Kappa and the other polymerases [Figure 4 and Table S4]. The total sequences with at least one error is also polymerase dependent with the majority of differences being between Kappa or Accuprime and the other polymerases [Table S3 & S4]. These differences in error rates were not due to polymerase dependent differences in base substitution rate [Figure S1]. Collectively, the results suggest that sequence error is dependent on polymerase, persists across the number of cycles used, and are not due to any bias towards a specific base subsitution.

***Prevalence of Chimeric Sequences are Polymerase Dependent and Correlate with the Number of OTUs:***  Based on the previous results, we examined whether chimeras also were dependent on polymerase and whether this could affect the number of OTUs. 
There is significant differences in the chimera prevalence based on polymerase at all the number of cycles used (P-value `r ifelse(max(filter(kruskal_chimera_prev)[, "bh"]) < 0.05, invisible("<"), invisible("check data"))` 0.05) [Table S3]. Differences in chimera prevalence between Platinum and all other polymerases accounted for the majority of these differences [Table S4]. Accuprime^TM^ had the lowest chimera prevalence of all polymerases regardless of whether chimera removal with VSEARCH was used [Figure 5A & 5B]. Additionally, there was a plateau in the total percent of chimeras that were removed that was similar for all polymerases [Figure 5C]. A positive correlation was observed between chimeric sequences and the number of OTUs for all polymerases [Figure 6]. This positive correlation was strongest for Accuprime^TM^, Platinum, and Phusion [Figure 6]. This data suggests that chimera prevalence depends on polymerase used and confirms that the number of OTUs is dependent on the prevalence of these chimeric sequences.

### 2018-05-29 submission/manuscript.tex
redo third section of results

\textbf{\emph{The number of OTUs are dependent on polymerase:}}
Differences in the range of the number of OTUs detected for fecal
samples is dependent on the polymerase used (e.g.~Accuprime at 35 cycles
= 84 - 106 versus Phusion at 35 cycles = 84 - 136) {[}Figure 1{]}.
Additionally, there is a trend for lower number of cycles (15-20) to
result in a reduced range in the number of OTUs versus higher number of
cycles (25, 30, and 35) for all polymerases (e.g.~Phusion at 15 cycles =
10 - 19 versus Phusion at 35 cycles = 84 - 136) {[}Figure 1{]}. There is
an overall difference in the number of OTUs detected within fecal
samples between polymerases at 35 cycles (F-stat \textgreater{} 16.35,
P-value = 9.7e-05) {[}Table S1{]}. Using a Tukey post-hoc test to
identify which polymerase groups were different from each other at 35
cycles, no difference was found (P-value \textgreater{} 0.05) {[}Table
S2{]}. The polymerase dependent difference in the range of the number of
OTUs also was observed in the mock community samples {[}Figure 2{]}. The
closest the polymerases came to the total of 8 OTUs created by the mock
reference 16S sequences was at 25 and 30 cycles {[}Figure 2{]}.
Regardless of if fecal or mock communities were used, the same
polymerases generated high and low number of OTUs and this was
consistent across the number of cycles used {[}Figure 1-2 \& Table
S1-S2{]}. In contrast to the results obtained with fecal samples,
differences between polymerases for the number of OTUs created were
observed as early as 20 cycles in the mock community (F-stat = 15.82,
P-value = 0.002) {[}Table S1{]}. Using a Tukey post-hoc test, the
majority of differences for the number of OTUs detected in the mock
community is largely due to Kappa and Platinum differences versus the
other polymerases across the different number of cycles {[}Table S2{]}.
Based on these observations in fecal and mock communities, it is clear
that using different polymerases result in a different total number of
OTUs within a sample.

@@

\textbf{\emph{Sequence Error is Dependent on both Polymerase and Cycle
Number:}} Differences in the median error rate varied by polymerase,
with Kappa being the highest {[}Figure 4 \& Table S3{]}. The differences
in the median error rate between the polymerases also is dependent on
the number of cycles, with Platinum and Kappa having the most
differences versus other polymerases {[}Figure 4 and Table S4{]}. The
total sequences with at least one error also is dependent on the number
of cycles {[}Table S3 \& S4{]}. These differences in sequences with at
least one error is mostly due to differences in
Accuprime\textsuperscript{TM} and Kappa versus the other polymerases
{[}Table S3 \& S4{]}. The differences in error rates were not due to
polymerase dependent differences in base substitution rate {[}Figure
S1{]}. These results suggest that sequence error is dependent on
polymerase and the number of cycles, but is not attributable to a
specific type. Based on these results, we examined whether chimeras also
were dependent on polymerase and whether this also could affect the
number of OTUs.

\textbf{\emph{Prevalence of Chimeric Sequences are Polymerase Dependent
and Correlate with the Number of OTUs:}} There is significant
differences in the chimera prevalence based on polymerase at all the
number of cycles used (P-value \textless{} 0.05) {[}Table S3{]}.
Differences in chimera prevalence between Platinum and all other
polymerases accounted for the majority of these differences {[}Table
S4{]}. Accuprime\textsuperscript{TM} had the lowest chimera prevalence
of all polymerases regardless of whether chimera removal with VSEARCH
was used {[}Figure 5A \& 5B{]}. Additionally, there was a plateau in the
total percent of chimeras that were removed that was similar for all
polymerases {[}Figure 5C{]}. A positive correlation was observed between
chimeric sequences and the number of OTUs for all polymerases {[}Figure
6{]}. This positive correlation was strongest for
Accuprime\textsuperscript{TM}, Platinum, and Phusion {[}Figure 6{]}.
This data suggests that chimera prevalence depends on polymerase used
and confirms that the number of OTUs is dependent on the prevalence of
these chimeric sequences.

>>>>

\textbf{\emph{The number of OTUs generated are dependent on polymerase
used:}} Differences in the range of the number of OTUs detected for
fecal samples is dependent on the polymerase used (e.g.~Accuprime at 35
cycles = 84 - 106 versus Phusion at 35 cycles = 84 - 136) {[}Figure
1{]}. Additionally, there is a trend for lower number of cycles (15-20)
to result in a reduced range in the number of OTUs versus higher number
of cycles (25, 30, and 35) for all polymerases (e.g.~Phusion at 15
cycles = 10 - 19 versus Phusion at 35 cycles = 84 - 136) {[}Figure 1{]}.
There is an overall difference in the number of OTUs detected within
fecal samples between polymerases at 35 cycles (F-stat \textgreater{}
16.35, P-value = 9.7e-05) {[}Table S1{]}. Using a Tukey post-hoc test to
identify which polymerase groups were different from each other at 35
cycles, no difference was found (P-value \textgreater{} 0.05) {[}Table
S2{]}. The polymerase dependent difference in the range of the number of
OTUs also was observed in the mock community samples {[}Figure 2{]}. The
closest the polymerases came to the total of 8 OTUs created by the mock
reference 16S sequences was at 25 and 30 cycles {[}Figure 2{]}.
Regardless of if fecal or mock communities were used, the same
polymerases generated high and low number of OTUs and this was
consistent across the number of cycles used {[}Figure 1-2 \& Table
S1-S2{]}. In contrast to the results obtained with fecal samples,
differences between polymerases for the number of OTUs created were
observed as early as 20 cycles in the mock community (F-stat = 15.82,
P-value = 0.002) {[}Table S1{]}. Using a Tukey post-hoc test, the
majority of differences for the number of OTUs detected in the mock
community is largely due to Kappa and Platinum differences versus the
other polymerases across the different number of cycles {[}Table S2{]}.
Based on these observations in fecal and mock communities, it is clear
that using different polymerases result in a different total number of
OTUs within a sample.

@@

\textbf{\emph{Sequence error varies by polymerase and is consistent
across the number of cycles used:}} The median error rate varied by
polymerase, with Kappa having the highest error rate of all the
polymerases across the number of cycles {[}Figure 4 \& Table S3{]}. The
majority of the differences across the number of cycles was between
Kappa and the other polymerases {[}Figure 4 and Table S4{]}. The total
sequences with at least one error is also polymerase dependent with the
majority of differences being between Kappa or Accuprime and the other
polymerases {[}Table S3 \& S4{]}. These differences in error rates were
not due to polymerase dependent differences in base substitution rate
{[}Figure S1{]}. Collectively, the results suggest that sequence error
is dependent on polymerase, persists across the number of cycles used,
and are not due to any bias towards a specific base subsitution.

\textbf{\emph{Prevalence of Chimeric Sequences are Polymerase Dependent
and Correlate with the Number of OTUs:}} Based on the previous results,
we examined whether chimeras also were dependent on polymerase and
whether this could affect the number of OTUs. There is significant
differences in the chimera prevalence based on polymerase at all the
number of cycles used (P-value \textless{} 0.05) {[}Table S3{]}.
Differences in chimera prevalence between Platinum and all other
polymerases accounted for the majority of these differences {[}Table
S4{]}. Accuprime\textsuperscript{TM} had the lowest chimera prevalence
of all polymerases regardless of whether chimera removal with VSEARCH
was used {[}Figure 5A \& 5B{]}. Additionally, there was a plateau in the
total percent of chimeras that were removed that was similar for all
polymerases {[}Figure 5C{]}. A positive correlation was observed between
chimeric sequences and the number of OTUs for all polymerases {[}Figure
6{]}. This positive correlation was strongest for
Accuprime\textsuperscript{TM}, Platinum, and Phusion {[}Figure 6{]}.
This data suggests that chimera prevalence depends on polymerase used
and confirms that the number of OTUs is dependent on the prevalence of
these chimeric sequences.
